---
title: "A2_Main_Quarto"
format: html
editor: visual
---

## Project Set Up:

```{r label = SetUp, echo = TRUE, warning = FALSE, message= FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
library(gridExtra)

theme_set(theme_minimal(base_size = 12))
set.seed(2026)
```

```{r label = LoadData, echo = TRUE}
dat <- read.table("HittingTheWall_2026.txt", header = TRUE)

# Make Bombed a Boolean:
dat$Bombed <- as.logical(dat$Bombed)

# Make Sex a Factor For Plotting Purposes Only:
dat$Sex_plot <- factor(dat$Sex, levels = c(0, 1), labels = c("Female", "Male"))
```

## a)

### Exploratory Data Analysis:

```{r label = EDAPlots, echo = TRUE}
# Average First-Half Speed vs. Bombed:
p1 <- ggplot(dat, aes(x = AveSpeed, y = as.numeric(Bombed))) +
  geom_jitter(height = 0.05, width = 0, alpha = 0.5) +
  geom_smooth(method = "loess", se = FALSE) +
  scale_y_continuous(
    breaks = c(0, 1),
    labels = c("FALSE", "TRUE")) +
  labs(y = "Bombed",
       x = "Average First-Half Speed (km/h)") +
  theme(
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size = 18),
    axis.text.x  = element_text(size = 14),
    axis.text.y  = element_text(size = 14)
  )
  

# Fluid Intake vs. Bombed:
p2 <- ggplot(dat, aes(x = Fluid, y = as.numeric(Bombed))) +
  geom_jitter(height = 0.05, width = 0, alpha = 0.5) +
  geom_smooth(method = "loess", se = FALSE) +
  scale_y_continuous(
    breaks = c(0, 1),
    labels = c("FALSE", "TRUE")) +
  labs(y = "Bombed",
       x = "Fluid Intake (l/h)") +
  theme(
    axis.title.x = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    axis.text.x  = element_text(size = 12),
    axis.text.y  = element_text(size = 12)
  )

# Carbohydrate Intake vs. Bombed:
p3 <- ggplot(dat, aes(x = Carbs, y = as.numeric(Bombed))) +
  geom_jitter(height = 0.05, width = 0, alpha = 0.5) +
  geom_smooth(method = "loess", se = FALSE) +
  scale_y_continuous(
    breaks = c(0, 1),
    labels = c("FALSE", "TRUE")) +
  labs(
    y = "Bombed",
    x = "Carbohydrate Intake (g/h)") +
  theme(
    axis.title.x = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    axis.text.x  = element_text(size = 12),
    axis.text.y  = element_text(size = 12)
  )

#  Sex vs. Bombed:
p4 <- ggplot(dat, aes(x = Sex_plot, y = as.numeric(Bombed))) +
  geom_jitter(height = 0.05, width = 0.1, alpha = 0.4) +
  stat_summary(fun = mean, geom = "point", size = 3, colour = "blue") +
  scale_y_continuous(
    breaks = c(0, 1),
    labels = c("FALSE", "TRUE")) +
  labs(y = "Bombed",
       x = "Sex") +
  theme(
    axis.title.x = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    axis.text.x  = element_text(size = 12),
    axis.text.y  = element_text(size = 12)
  )

p1
p2 
p3
p4

```

## b)

## c)

```{r label = SoftmaxFunction, echo = TRUE}
# Definition of Softmax Activation Function for Input Matrix Z:
softmax <- function(Z) {
  # Calculation of Shifted Matrix by Max Column-Wise Value for Numerical Stability:
  Z_shift <- sweep(Z, 2, apply(Z, 2, max), FUN = "-")
  # Element-Wise Exponentiation of Shifted Matrix:
  exp_Z <- exp(Z_shift)
  # Normalisation of Each Column, Returning Class Probabilities:
  exp_Z / matrix(colSums(exp_Z), nrow = nrow(Z), ncol = ncol(Z), byrow = TRUE)
}
```

## d)

## e)

### Create Input and Output Matrices:

```{r label = MatricesCreation, echo = TRUE}
# Make Bombed Variable Boolean:
dat$Bombed <- as.logical(dat$Bombed)

# Keep a Numeric Version for Modelling:
dat$Sex_num <- ifelse(dat$Sex == "Male" | dat$Sex == 1, 1, 0)

# Creation of Xu = Demographic/Performance Inputs:
Xu <- t(as.matrix(dat[, c("AveSpeed", "Sex_num")]))
storage.mode(Xu) <- "double"

# Creation of Xv = Nutrition Inputs:
Xv <- t(as.matrix(dat[, c("Fluid", "Carbs")]))
storage.mode(Xv) <- "double"

# Creation of Two-Output Response Matrix:
Y <- rbind(Bombed_FALSE = as.numeric(!dat$Bombed), Bombed_TRUE  = as.numeric(dat$Bombed))
```

## f)

## g)

### Feed Forward & Objective Function

```{r label = SplitBrainNetworkFunctions, echo = TRUE}
# Function to Unpack Parameter Vector into Split-Brain Network Weights and Biases:
unpack_split_params <- function(theta, pu, pv, m, q) {

  # Index for Position Tracking in Parameter Vector:
  idx <- 1
  
  # Helper Function to Extract Next k Parameters from Theta:
  take <- function(k) {
    out <- theta[idx:(idx + k - 1)]
    idx <<- idx + k
    out
  }
  
  # Extraction of First-Layer Weights and Biases for u-Hemisphere:
  W1u <- matrix(take(m * pu), nrow = m, ncol = pu)
  b1u <- matrix(take(m), nrow = m, ncol = 1)
  
  # Extraction of First-Layer Weights and Biases for v-Hemisphere:
  W1v <- matrix(take(m * pv), nrow = m, ncol = pv)
  b1v <- matrix(take(m), nrow = m, ncol = 1)
  
  # Extraction of Second-Layer Weights and Biases for u-Hemisphere:
  W2u <- matrix(take(m), nrow = 1, ncol = m)
  b2u <- matrix(take(1), nrow = 1, ncol = 1)
  
  # Extraction of Second-Layer Weights and Biases for v-Hemisphere:
  W2v <- matrix(take(m), nrow = 1, ncol = m)
  b2v <- matrix(take(1), nrow = 1, ncol = 1)
  
  # Extraction of Output-Layer Weights and Biases:
  W3 <- matrix(take(q * 2), nrow = q, ncol = 2)
  b3 <- matrix(take(q), nrow = q, ncol = 1)
  
  # Return All Unpacked Parameters as Named List:
  list(
    W1u = W1u, b1u = b1u,
    W1v = W1v, b1v = b1v,
    W2u = W2u, b2u = b2u,
    W2v = W2v, b2v = b2v,
    W3 = W3, b3 = b3)
}

# Function to Count Total Number of Parameters in Split-Brain Network:
n_params_split <- function(pu, pv, m, q) {
  m*(pu + pv + 2) + 2*q + 2*m + 2 + q
}

# Function to Randomly Initialise Split-Brain Network Parameters:
init_split_params <- function(pu, pv, m, q, sd = 0.1) {
  rnorm(n_params_split(pu, pv, m, q), mean = 0, sd = sd)
}

# Feed-Forward Function for Split-Brain Network with Objective Function Evaluation:
forward_split <- function(theta, Xu, Xv, Y = NULL, m = 4, nu = 0) {

  # Extraction of Input and Output Dimensions from Supplied Matrices:
  pu <- nrow(Xu)
  pv <- nrow(Xv)
  q  <- 2
  N  <- ncol(Xu)
  
  # Unpacking Parameter Vector into Structured Weight and Bias Matrices:
  pars <- unpack_split_params(theta, pu, pv, m, q)
  
  # Creation of Row Vector of Ones to "Broadcast" Bias Terms Across Observations:
  oneN <- matrix(1, nrow = 1, ncol = N)
  
  # Computation of First Hidden-Layer Linear Combinations & Activations for u-Hemisphere & v-Hemisphere:
  Z1u <- pars$W1u %*% Xu + pars$b1u %*% oneN
  A1u <- tanh(Z1u)
  Z1v <- pars$W1v %*% Xv + pars$b1v %*% oneN
  A1v <- tanh(Z1v)
  
  # Computation of Second Hidden-Layer Linear Combinations & Activation for u-Hemisphere & v-Hemisphere:
  Z2u <- pars$W2u %*% A1u + pars$b2u %*% oneN
  A2u <- tanh(Z2u)
  Z2v <- pars$W2v %*% A1v + pars$b2v %*% oneN
  A2v <- tanh(Z2v)
  
  # Row Binding the Two Hemisphere Activations for Input to Output Layer:
  A2 <- rbind(A2u, A2v)
  
  # Computation of Output-Layer Linear Combinations and Softmax Probabilities:
  Z3 <- pars$W3 %*% A2 + pars$b3 %*% oneN
  P  <- softmax(Z3)
  
  # Storage of Feed Forward Outputs in a List:
  out <- list(
    P = P,
    Z1u = Z1u, A1u = A1u,
    Z1v = Z1v, A1v = A1v,
    Z2u = Z2u, A2u = A2u,
    Z2v = Z2v, A2v = A2v,
    Z3 = Z3)
  
  # Check for Response Matrix so that Objective Function can be Evaluated:
  if (!is.null(Y)) {

    # Definition of Small Constant to Avoid log(0):
    eps <- 1e-12

    # Computation of Cross-Entropy Component of the Objective Function
    ce <- -mean(colSums(Y * log(P + eps)))
    
    # Computation of L2 Regularisation Penalty on Weight Matrices Component:
    reg <- (nu / 2) * (sum(pars$W1u^2) + sum(pars$W1v^2) + sum(pars$W2u^2) + sum(pars$W2v^2) + sum(pars$W3^2))
    
    # Storage of Total Objective Function & Spearate Compenents:
    out$loss <- ce + reg
    out$ce   <- ce
    out$reg  <- reg
  }
  
  # Return of Feed-Forward Results:
  out
}

# Function to Convert Predicted Probabilities Into Class Predictions:
predict_split_class <- function(theta, Xu, Xv, m = 4) {
  P <- forward_split(theta, Xu, Xv, m = m)$P
  max.col(t(P)) - 1
}

# Function to Compute Misclassification Rate:
misclass_rate <- function(pred_class, Y) {
  truth <- max.col(t(Y)) - 1
  mean(pred_class != truth)
}
```

Fitting Function:

```{r label = FitSplitFunction, echo = TRUE}
fit_split_network <- function(Xu_train, Xv_train, Y_train, m = 4, nu = 0, maxit = 1000, n_starts = 3, seed = 2026) {
  
  # Setting Seed for Reproducibility:
  set.seed(seed)
  
  # Determining Input and Output Dimensions:
  pu <- nrow(Xu_train)
  pv <- nrow(Xv_train)
  q  <- nrow(Y_train)
  
  # Initialising Objects to Track the Best Fit Across Random Starts:
  best_fit <- NULL
  best_val <- Inf
  
  # Running Multiple Starts to Reduce the Risk of Poor Local Minima:
  for (s in 1:n_starts) {
    
    # Generating a Random Initial Parameter Vector:
    theta0 <- init_split_params(pu, pv, m, q, sd = 0.1)
    
    # Optimising the Split-Brain Network Objective Function:
    fit <- optim(
      par = theta0,
      fn = function(th) forward_split(th, Xu_train, Xv_train, Y_train, m = m, nu = nu)$loss,
      method = "BFGS",
      control = list(maxit = maxit, reltol = 1e-8)
    )
    
    # Updating the Stored Best Fit if the Current Start Improves the Objective Value:
    if (fit$value < best_val) {
      best_val <- fit$value
      best_fit <- fit
    }
  }
  
  # Returning the Best Optimisation Result Across All Starts:
  best_fit
}
```

Since the response is binary and the network outputs class probabilities, the chosen objective function is the multinomial cross-entropy loss. To control model complexity, an L2 penalty was added on the weight matrices, with regularisation level denoted by v. Bias terms were left unpenalised.

## h)

### Validation Analysis for m = 4:

Train-Validation Split

```{r label = SplitData, echo = TRUE}
# Set Seed for Reproducibility:
set.seed(2026)
# Calculate the Number of Observations:
n <- nrow(dat)
# Randomly Select 80% of the Data for the Training Set:
train_id <- sample(seq_len(n), size = floor(0.8 * n))
# Use the Remaining 20% as the Validation Set:
valid_id <- setdiff(seq_len(n), train_id)

# Subset Input and Response Matrices for the Training Set:
Xu_train <- Xu[, train_id, drop = FALSE]    #The u-hemisphere inputs.
Xv_train <- Xv[, train_id, drop = FALSE]    #The v-hemisphere inputs.
Y_train  <- Y[, train_id, drop = FALSE]   #The responses.

# Subset Input and Response Matrices for the Validation Set:
Xu_valid <- Xu[, valid_id, drop = FALSE]    #The u-hemisphere inputs.
Xv_valid <- Xv[, valid_id, drop = FALSE]    #The v-hemisphere inputs.
Y_valid  <- Y[, valid_id, drop = FALSE]   #The responses.
```

Validation Loop Over Regularisation Values

```{r label = ValidationGrid, echo = TRUE}
# Create a Grid of Nu (Regularisation) Values on a Logarithmic Scale:
nu_grid <- exp(seq(-8,0,length = 20)) 

# If the Tuning has Already Been Run, Used Saved Results:
if (file.exists("ValidationLoopResults.Rdata"))
{
  load("ValidationLoopResults.Rdata")
} else 
{
  # Create a Data Frame to store the different results for each nu:
  val_results <- data.frame(
    nu = nu_grid,
    train_loss = NA_real_, #Training loss (penalised objective). 
    train_ce = NA_real_, #Training cross-entropy. 
    valid_ce = NA_real_, #Validation cross-entropy. 
    train_err = NA_real_, #Training misclassification error. 
    valid_err = NA_real_ #Validation misclassification error.
    )
  
  # Store Fitted Models for Each Nu:
  split_fits <- vector("list", length(nu_grid))
  
  # Loop Over Our Nu Grid:
  for (i in seq_along(nu_grid)) 
  {
    # Fit the Split-Brain Network Using the Current Nu Value:
    fit_i <- fit_split_network(
      Xu_train, Xv_train, Y_train,
      m = 4, nu = nu_grid[i],
      maxit = 800, n_starts = 8, seed = 2026 + i)
    
    # Store the Model:
    split_fits[[i]] <- fit_i
    
    # Evaluate the Model Performance on the Training and Validation Sets:
    tr_eval <- forward_split(fit_i$par, Xu_train, Xv_train, Y_train, m = 4, nu = nu_grid[i])
    va_eval <- forward_split(fit_i$par, Xu_valid, Xv_valid, Y_valid, m = 4, nu = nu_grid[i])
    
    # Make Predicted Class Labels for Training and Validation Sets:
    tr_pred <- predict_split_class(fit_i$par, Xu_train, Xv_train, m = 4)
    va_pred <- predict_split_class(fit_i$par, Xu_valid, Xv_valid, m = 4)
    
    # Store the Results:
    val_results$train_loss[i] <- tr_eval$loss  #Training loss (penalised objective). 
    val_results$train_ce[i]   <- tr_eval$ce  #Training cross-entropy.  
    val_results$valid_ce[i]   <- va_eval$ce  #Validation cross-entropy.
    val_results$train_err[i]  <- misclass_rate(tr_pred, Y_train)  #Training misclassification error. 
    val_results$valid_err[i]  <- misclass_rate(va_pred, Y_valid)  #Validation misclassification error. 
  }
  # Save the Results for Future Use:
  save(val_results, split_fits, file = "ValidationLoopResults.Rdata")
}

# Display the Results:
val_results
```

Select Best Regularisation Level

```{r label = ChooseBestNu, echo = TRUE}
# Indentifying Index of the Best Nu Value (the one that minimises the validation cross-entropy):
best_idx <- which.min(val_results$valid_ce)
# Extracting this Corresponding Optimal Nu Value:
best_nu  <- val_results$nu[best_idx]
# Extracting this Corresponding Fitted Model:
best_fit_split <- split_fits[[best_idx]]

# Displaying this Nu Value and its Corresponding Performance Metrics:
best_nu
val_results[best_idx, ]
```

Validation Plot

```{r label = ValidationPlot, echo = TRUE}
#| fig-width: 6
#| fig-height: 6

# Plotting Validation Cross-Entropy Against the Regularisation Parameter Nu: 
ggplot(val_results, aes(x = nu, y = valid_ce)) +
  geom_line() +
  geom_point(data = val_results[best_idx, ], size = 4, colour = "red") +
  geom_point() +
  scale_x_log10() +
  labs(x = expression(nu), y = "Validation cross-entropy")
```

Figure 5 plots the validation cross-entropy against the regularisation parameter $\nu$. The curve initially decreases and reaches a minimum after which it increases sharply and eventually plateaus at higher values of $\nu$. The minimum value validation cross-entropy value occurs at $\hat{\nu} = 0.001186388$, and this is therefore selected as the optimal level of regularisation. This point is highlighted on the plot.

This chosen $\nu$ value achieves a good balance between model fit and complexity, as it minimises the validation cross-entropy.

## i)

### Response curves over Fluid, Carbs, and Sex for a 12 km/h Runner

Build Prediction Grid

```{r label = ResponseGrid, echo = TRUE}
# Create Sequences Over Suitable Range of Fluid and Carbs Values:
fluid_seq <- seq(min(dat$Fluid), max(dat$Fluid), length.out = 100)
carb_seq  <- seq(min(dat$Carbs), max(dat$Carbs), length.out = 100)

# Create a Grid of all Combinations of Fluid, Carbs and Sex:
grid_pred <- expand.grid(Fluid = fluid_seq, Carbs = carb_seq, Sex = c("Female", "Male"))

# Fix the AveSpeed for Each Combination as 12km/h:
grid_pred$AveSpeed <- 12

# Convert Sex into a Numeric Variable:
grid_pred$SexNum <- ifelse(grid_pred$Sex == "Male", 1, 0)

# Create Input Matrices as Needed by Split-Brain Network:
Xu_grid <- t(as.matrix(grid_pred[, c("AveSpeed", "SexNum")]))
Xv_grid <- t(as.matrix(grid_pred[, c("Fluid", "Carbs")]))

# Evaluate Fitted Model for Each Input in the Grid:
P_grid <- forward_split(best_fit_split$par, Xu_grid, Xv_grid, m = 4)$P

# Get the Predicted Probabilities of Bombing for Each Input in the Grid:
grid_pred$ProbBombed <- as.numeric(P_grid[2, ])
```

Surface-Style Heatmaps

```{r label = ResponseHeatmaps, echo = TRUE}
#| fig-width: 8
#| fig-height: 4

# Response Surface for 12km/h Target - Predicted Bombing Probability over Fluid and Carbs Faceted by Sex:
ggplot(grid_pred, aes(x = Fluid, y = Carbs, fill = ProbBombed)) +
  geom_tile() +
  facet_wrap(~ Sex) +
  labs(fill = "Pr(Bombed)") +
  theme(
    axis.title.x = element_text(size = 22),
    axis.title.y = element_text(size = 22),
    axis.text.x  = element_text(size = 17),
    axis.text.y  = element_text(size = 17),
    strip.text = element_text(size = 17))
```

## j)

The response curves show how the estimated probability of bombing varies with fluid intake and carbohydrate intake for runners targeting 12 km/h over the first half of the marathon. In both panels, it is clear that there is an optimal region of low probability of bombing (dark area), which is surrounded by a higher risk area as one leaves this zone (lighter area).

The lowest probability of bombing happens when runners have a moderate fluid intake (around 1 to 1.6 liters per hour), and a moderate to high carbohydrate intake (around 60 to 80 grams per hour). Moving away from this results in a far higher probability of bombing.

This means that for a runner targeting a 12km/h pace over the first half of the race, it is optimal for them to be taking in between 1 and 1.6 liters per hour of fluids, and between 60 and 80 grams per hour of carbohydrates. In particular, runners should try to avoid having very low carbohydrate intake, as this increases the probability of bombing a lot, even if fluid intake is sufficient.

When comparing the two panels, the shape of the response surfaces is the same for each sex, which indicates that the relationship between bombing and carbohydrate and fluid intake is consistent between sex. However, the region of low probability of bombing for males is slightly smaller than for females, which indicates that males have a slightly smaller window of fluid and carbohydrate intake which minimises the probability of bombing. This does not, however, change the optimal fueling strategy for either sex.

## k)

### Empirical Advantage of the Split-Brain Network:

Fit a Comparable Standard Dense Network

```{r label = StandardNetworkFunctions, echo = TRUE}
# Combine the Split Matrices into One Input Matrix:
X_full <- rbind(Xu, Xv)

# Function to Unpack Parameter Vector into Standard Network Weights and Biases:
unpack_standard_params <- function(theta, p, m1, m2, q) {
  
  # Index for Position Tracking in Parameter Vector:
  idx <- 1
  
  # Helper Function to Extract Next k Parameters from Theta:
  take <- function(k) {
    out <- theta[idx:(idx + k - 1)]
    idx <<- idx + k
    out
  }
  
  # First Hidden Layer Parameters:
  W1 <- matrix(take(m1 * p), nrow = m1, ncol = p)
  b1 <- matrix(take(m1), nrow = m1, ncol = 1)
  
  # Second Hidden Layer Parameters:
  W2 <- matrix(take(m2 * m1), nrow = m2, ncol = m1)
  b2 <- matrix(take(m2), nrow = m2, ncol = 1)
  
  # Third Hidden Layer Parameters:
  W3 <- matrix(take(q * m2), nrow = q, ncol = m2)
  b3 <- matrix(take(q), nrow = q, ncol = 1)
  
  list(W1 = W1, b1 = b1, W2 = W2, b2 = b2, W3 = W3, b3 = b3)
}

#Function to Determine the Number of Parameters
n_params_standard <- function(p, m1, m2, q) {
  m1*p + m1 + m2*m1 + m2 + q*m2 + q
}

# Initialising the Parameters Randomly:
init_standard_params <- function(p, m1, m2, q, sd = 0.1) {
  rnorm(n_params_standard(p, m1, m2, q), mean = 0, sd = sd)
}

# Feed-Forward Function for Standard Network with Objective Function Evaluation:
forward_standard <- function(theta, X, Y = NULL, m1 = 8, m2 = 2, nu = 0) {
  p <- nrow(X)
  q <- 2
  N <- ncol(X)
  
  # Unpacking Parameter Vector into Structured Weight and Bias Matrices:
  pars <- unpack_standard_params(theta, p, m1, m2, q)
  
  # Creation of Row Vector of Ones to "Broadcast" Bias Terms Across Observations:
  oneN <- matrix(1, nrow = 1, ncol = N)
  
  # First Hidden Layer:
  Z1 <- pars$W1 %*% X + pars$b1 %*% oneN
  A1 <- tanh(Z1)
  
  # Second Hidden Layer:
  Z2 <- pars$W2 %*% A1 + pars$b2 %*% oneN
  A2 <- tanh(Z2)
  
  # Output Layer:
  Z3 <- pars$W3 %*% A2 + pars$b3 %*% oneN
  P <- softmax(Z3)
  
  out <- list(P = P)
  
  # Check for Response Matrix so that Objective Function can be Evaluated:
  if (!is.null(Y)) {
    # Definition of Small Constant to Avoid log(0):
    eps <- 1e-12
    # Computation of Cross-Entropy Component of the Objective Function
    ce <- -mean(colSums(Y * log(P + eps)))
    # Computation of L2 Regularisation Penalty on Weight Matrices Component:
    reg <- (nu / 2) * (sum(pars$W1^2) + sum(pars$W2^2) + sum(pars$W3^2))
    # Storage of Total Objective Function & Spearate Compenents:
    out$loss <- ce + reg
    out$ce <- ce
    out$reg <- reg
  }
  
  # Return Feed-Forward Result:
  return(out)
}

# Function to Fit the Standard NN Using Optim Function:
fit_standard_network <- function(X_train, Y_train, m1 = 8, m2 = 2, nu = 0, maxit = 800, n_starts = 8, seed = 2026) {
  set.seed(seed)
  
  p <- nrow(X_train)
  q <- nrow(Y_train)
  
  best_fit <- NULL
  best_val <- Inf
  
  # Multiple Starts to Avoid Poor Local Minimums:
  for (s in 1:n_starts) {
    theta0 <- init_standard_params(p, m1, m2, q, sd = 0.1)
    
    fit <- optim(
      par = theta0,
      fn = function(th) forward_standard(th, X_train, Y_train, m1 = m1, m2 = m2, nu = nu)$loss,
      method = "BFGS",
      control = list(maxit = maxit, reltol = 1e-8))
    
    # Store the Best Solution:
    if (fit$value < best_val) {
      best_val <- fit$value
      best_fit <- fit
    }
  }
  best_fit
}

# Generate Predicted Class Labels from the Standard NN:
predict_standard_class <- function(theta, X, m1 = 8, m2 = 2) {
  P <- forward_standard(theta, X, m1 = m1, m2 = m2)$P
  max.col(t(P)) - 1
}
```

Compare Split-Brain and Standard Networks

```{r label = CompareModels, echo = TRUE}
# Subset Data into Training and Validation Sets:
X_train <- X_full[, train_id, drop = FALSE]
X_valid <- X_full[, valid_id, drop = FALSE]

if (file.exists("NetworkComparisonObjects.RData"))
{
  load("NetworkComparisonObjects.RData")
} else 
{
  # Use Same best_nu for a Fair First Comparison and Fit a Standard NN:
  fit_standard_bestnu <- fit_standard_network(
    X_train, Y_train,
    m1 = 8, m2 = 2, nu = best_nu,
    maxit = 800, n_starts = 8, seed = 999)
  
  # Determine Validation Cross Entropy for Each Model:
  split_valid_ce <- forward_split(best_fit_split$par, Xu_valid, Xv_valid, Y_valid, m = 4, nu = best_nu)$ce
  std_valid_ce   <- forward_standard(fit_standard_bestnu$par, X_valid, Y_valid, m1 = 8, m2 = 2, nu = best_nu)$ce
  
  # Determine Validation Misclassification for Each Model:
  split_valid_err <- misclass_rate(
    predict_split_class(best_fit_split$par, Xu_valid, Xv_valid, m = 4), Y_valid)
  
  std_valid_err <- misclass_rate(
    predict_standard_class(fit_standard_bestnu$par, X_valid, m1 = 8, m2 = 2), Y_valid)
  
  #Save Results:
  save(fit_standard_bestnu, split_valid_ce, std_valid_ce, split_valid_err, std_valid_err, file = "NetworkComparisonObjects.RData")
}

# Create Performance Comparison Results Table:
comparison_tbl <- data.frame(
  Model = c("Split-Brain", "Standard Network"),
  Validation_CE = c(split_valid_ce, std_valid_ce),
  Validation_Misclassification = c(split_valid_err, std_valid_err))

# Convert Table to Latex:
cat(knitr::kable(
        comparison_tbl,
        format = "latex",
        booktabs = TRUE,
        caption = "Comparison of split-brain vs. standard dense NN performance.",
        col.names = c("Model", "Validation_CE", "Validation_Misclassification"),
        align = c("l", "c", "c")))
```

To test whether the split-brain network provides an empirical advantage, a standard neural network was fit using the same training and validation data, at the same value of $\hat{\nu}$, and yielded the following results:

The standard neural network has a far lower validation cross-entropy of $0.053$ vs $0.107$ for the split brain network. The standard neural network also has a lower misclassification error of $0.014$ vs $0.056$ for the split brain network.

This provides clear empirical evidence that the split-brain network does not improve the predictive performance of this problem.
